---
title: "Take Home Ex 4"
author: "Aruiana"
date: 21 Feb 2023
date-modified: "`r Sys.Date()`"
format: html
editor: visual
---

# 1. The Task

To uncover the impact of COVID-1 and global economic and political dynamics happening in 2022 on Singapore bi-lateral trade.

For the purpose of this study, the focus in on **January 2020 to Dec 2022**.

# 2. Data Preparation

## Step 1: Load Packages

```{r}
#| code-fold: true
#| code-summary: "Show"
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table, CGPfunctions, ggHoriPlot, tidyverse, datawizard)
```

## Step 2: Import Data

```{r}
#| code-fold: true
#| code-summary: "Show"
#import data
i <- data_transpose(read_xlsx("data/outputFile.xlsx", sheet = "T1", range = "A10:AK129" ), rownames = "imtime", colnames = TRUE)

e <- data_transpose(read_xlsx("data/outputFile.xlsx", sheet = "T2", range = "A10:AK101" ), rownames = "extime", colnames = TRUE)
```

## Step3: Clean Data (Import)

```{r}
#| code-fold: true
#| code-summary: "Show"

#convert data to raw figures
importdata <- i %>%
  mutate_at(vars(contains('Thousand Dollars')), ~ (. *1000)) %>%      
  mutate_at(vars(contains('Million Dollars')), ~ (. *1000000)) %>%
  separate(imtime, into = c("year", "month"))


#remove legend
colnames(importdata) <- gsub(" (Thousand Dollars)", "", colnames(importdata), fixed = TRUE)
colnames(importdata) <- gsub(" (Million Dollars)", "", colnames(importdata), fixed = TRUE)
colnames(importdata) <-gsub('Vietnam, Socialist Republic Of', 'Vietnam',gsub('Germany, Federal Republic Of', 'Germany', gsub('Republic of Korea', 'Korea', colnames(importdata))))

#convert date to numbers
importdata$month <- gsub('Dec', '12', gsub('Nov','11', gsub('Oct','10',gsub('Sep','09',gsub('Aug','08', gsub('Jul','07', gsub('Jun', '06', gsub('May', '05',gsub('Apr','04', gsub('Mar','03', gsub('Feb','02', gsub('Jan','01',importdata$month)))))))))))) %>% as.numeric(importdata$month)

importdata$year <- as.numeric(importdata$year)


#extract only country data and non-zero data
imd <- importdata [,!names(importdata) %in% c("Total Merchandise Imports", "Europe", "America", "Asia", "Europe", "Oceania", "Africa", "European Union","Commonwealth Of Independent States")]
import_use <- imd[, colSums(imd) != 0]


import_use$time = as.Date(paste(import_use$year,"-",import_use$month,"-01",sep = ""))

imd2 <- import_use [,!names(import_use) %in% c("year", "month")]

impexp <- imd2 %>% pivot_longer(!time, names_to = "Country", values_to = "Import")


```

## Step4: Clean Data (Export)

```{r}
#| code-fold: true
#| code-summary: "Show"

#repeat data cleaning for export

#convert data to raw figures
exportdata <- e %>%
  mutate_at(vars(contains('Thousand Dollars')), ~ (. *1000)) %>%      
  mutate_at(vars(contains('Million Dollars')), ~ (. *1000000)) %>%
  separate(extime, into = c("year", "month"))


#remove legend
colnames(exportdata) <- gsub(" (Thousand Dollars)", "", colnames(exportdata), fixed = TRUE)
colnames(exportdata) <- gsub(" (Million Dollars)", "", colnames(exportdata), fixed = TRUE)
colnames(exportdata) <-gsub('Vietnam, Socialist Republic Of', 'Vietnam',gsub('Germany, Federal Republic Of', 'Germany',gsub('Republic of Korea', 'Korea', colnames(exportdata))))

#convert date to numbers
exportdata$month <- gsub('Dec', '12', gsub('Nov','11', gsub('Oct','10',gsub('Sep','09',gsub('Aug','08', gsub('Jul','07', gsub('Jun', '06', gsub('May', '05',gsub('Apr','04', gsub('Mar','03', gsub('Feb','02', gsub('Jan','01',exportdata$month)))))))))))) %>% as.numeric(exportdata$month)

exportdata$year <- as.numeric(exportdata$year)


#extract only country data and non-zero data
exd <- exportdata [,!names(exportdata) %in% c("Total Merchandise Exports", "Europe", "America", "Asia", "Europe", "Oceania", "Africa", "European Union", "Commonwealth Of Independent States")]
export_use <- exd[, colSums(exd) != 0]


export_use$time = as.Date(paste(export_use$year,"-",export_use$month,"-01",sep = ""))

exd2 <- export_use [,!names(export_use) %in% c("year", "month")]

impexp2 <- exd2 %>% pivot_longer(!time, names_to = "Country", values_to = "Export")

```

## Step 5: Combine Data

```{r}
#| code-fold: true
#| code-summary: "Show"
impexp_use <- merge(x=impexp,y=impexp2,by.x=c("time","Country"),by.y=c("time","Country"),all.x = TRUE,all.y = TRUE)
  
impexp_use[is.na(impexp_use)] = 0

impexp_use$nett <- impexp_use$Export - impexp_use$Import
```

## Step 6: List top import countries

```{r}
#| code-fold: true
#| code-summary: "Show"
ImTop_2020 <- impexp_use %>%
  filter(time < '2021-01-01') %>%
  group_by(Country) %>%
  summarise(Imp2010 = sum(Import)) %>%
  top_n(n=10, wt=Imp2010) %>%
  arrange(desc(Imp2010))

ImTop_2021 <- impexp_use %>%
  filter(time >='2021-01-01' & time <='2021-12-31')  %>%
  group_by(Country) %>%
  summarise(Imp2110 = sum(Import)) %>%
  top_n(n=10, wt=Imp2110) %>%
  arrange(desc(Imp2110))

ImTop_2022 <- impexp_use %>%
  filter(time > '2022-01-01') %>%
  group_by(Country) %>%
  summarise(Imp2210 = sum(Import)) %>%
  top_n(n=10, wt=Imp2210) %>%
  arrange(desc(Imp2210))

AllImtop10 <- c(ImTop_2020$Country, ImTop_2021$Country, ImTop_2022$Country)
Top10Import <- unique(AllImtop10)


ImTop10 <- impexp_use %>%
 filter(Country %in% Top10Import)

```

## Step 6: List top export countries

```{r}
#| code-fold: true
#| code-summary: "Show"
ExTop_2020 <- impexp_use %>%
  filter(time < '2021-01-01') %>%
  group_by(Country) %>%
  summarise(Exp2010 = sum(Export)) %>%
  top_n(n=10, wt=Exp2010) %>%
  arrange(desc(Exp2010))

ExTop_2021 <- impexp_use %>%
  filter(time >='2021-01-01' & time <='2021-12-31')  %>%
  group_by(Country) %>%
  summarise(Exp2110 = sum(Export)) %>%
  top_n(n=10, wt=Exp2110) %>%
  arrange(desc(Exp2110))

ExTop_2022 <- impexp_use %>%
  filter(time > '2022-01-01') %>%
  group_by(Country) %>%
  summarise(Exp2210 = sum(Export)) %>%
  top_n(n=10, wt=Exp2210) %>%
  arrange(desc(Exp2210))

AllExtop10 <- c(ExTop_2020$Country, ExTop_2021$Country, ExTop_2022$Country)
Top10Export <- unique(AllExtop10)

ExTop10 <- impexp_use %>%
 filter(Country %in% Top10Export)

Alltop10 <- c(ExTop_2020$Country, ExTop_2021$Country, ExTop_2022$Country, ImTop_2020$Country, ImTop_2021$Country, ImTop_2022$Country)
Top10 <- unique(Alltop10)

IETop10 <- impexp_use %>%
 filter(Country %in% Top10)

```


#cycleplot

```{r}
#| code-fold: true
#| code-summary: "Show"
options(scipen = 999)
hline.data <- export_use %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(Belgium))

xbrks <- seq(2020,2022,1)
ggplot() + 
  geom_line(data=export_use,
            aes(x=year, 
                y=Belgium, 
                group=month), 
            colour="black") +   
  geom_hline(aes(yintercept= avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) + 
  labs(axis.text.x = element_blank(),
       title = "Export to Belgium") +
  xlab("Year") +
  ylab("$") + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) + scale_x_continuous(breaks=xbrks)

```

```{r}

cycle_ard <- function(country){
  options(scipen = 999)
  hline.data <- export_use %>% 
    group_by(month) %>%
    summarise(avgvalue = mean(country))
  
  xbrks <- seq(2020,2022,1)
  ggplot() + 
    geom_line(data=export_use,
              aes(x=year, 
                  y=country, 
                  group=month), 
              colour="black") +   
    geom_hline(aes(yintercept= avgvalue), 
               data=hline.data, 
               linetype=6, 
               colour="red", 
               size=0.5) + 
    facet_grid(~month) + 
    labs(axis.text.x = element_blank(),
         title = paste0("Export to ",country)) +
    xlab("Year") +
    ylab("$") + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) + 
    scale_x_continuous(breaks=xbrks)
}

cycle_ard("Belgium")

```

#slopegraph

```{r}

IETop10 %>% 
  mutate(Year = factor(time)) %>%
  filter(Year %in% c("2020-01-01","2021-01-01","2022-12-01")) %>%
  newggslopegraph(Year, Import, Country,
                Title = "Import by Singapore to Top countries",
                SubTitle = "2020 to 2022", Caption = "Singapore Imports as at date")

IETop10 %>% 
  mutate(Year = factor(time)) %>%
  filter(Year %in% c("2020-01-01","2021-01-01","2022-12-01")) %>%
  newggslopegraph(Year, Export, Country,
                Title = "Export by Singapore to Top countries",
                SubTitle = "2020 to 2022", Caption = "Singapore Exports as at date")


IETop10 %>% 
  mutate(Year = factor(time)) %>%
  filter(Year %in% c("2020-01-01","2021-01-01","2022-12-01")) %>%
  newggslopegraph(Year, nett, Country,
                Title = "Nett Import/Export by Singapore to Top countries",
                SubTitle = "2020 to 2022", Caption = "Nett import/export as at date")

```
